### 2013-09-03 Tex/chapters/model.tex
Formulierungs und Inhaltskorrekturen

    The most common definition of the Ising model, to which will be referred
    as the \emph{square lattice Ising model}, is a square lattice with edge length \(L\) and
    \(N=L^2\) sites. Each site has a magnetic moment, which is called the spin. Each

@@

        H = - \sum_{\avg{i,j}}J_{ij}s_{i}s_{j} - \sum_i h_i s_i.

@@

    the coupling constant. If \(J_{ij} > 0 \ \forall i,j\) the model resembles a ferromagnet.
    For the square lattice Ising model \(J_{ij} = 1 \ \forall i,j\) applies.
    \(h_i\) denotes the outer magnetic field at the position of
    site \(i\). For the square lattice Ising model \(h_i=0 \ \forall i\)
    applies.\\

@@

    thesis is that the sites of the square lattice are displaced.

@@

    \(\sigma\) will also be called \emph{disorder parameter} in the following.
    Most sites will only have one nearest neighbor after the

@@

    would destroy the two dimensional character of the model and the results
    would not be comparable with known results for the two dimensional
    Ising model.

@@

    so that the lattice represents a proximity graph. Edges
    of the used proximity graph do not cross each other, hence the 2D
    character of the lattice is preserved. The coupling constant \(J\) gets
    identified with edge weights. The weight of an edge \(E_{ij} =
    J_{ij} = \exp (\alpha (1-d_{ij}))\) where \(d_{ij}\) is the Euclidean
    distance between the nodes \(i\) and \(j\). The free parameter
    \(\alpha\) is set to \(\alpha = 0.5\) inspired by \cite{Lima2000}.

@@

            Note that the lattice on this torus is not a square but has
            a height to width ratio of 1:4. At 1:1 the torus would cut

@@

            %Also the shades are of course only a guide to the eye.

@@

    already studied on a Delaunay triangulation. \cite{Janke1994} examines
    this for constant \(J\) and \cite{Lima2000} for \(J_{ij} = e^{-\alpha d_{ij}}\).

@@

    exponents.

\subsection{Gabriel- and Relative Neighborhood Graph}
\label{ssec:graphtypes}
    A graph \(G(V,E)\) is a set of nodes \(V\) and edges \(E\). In the
    scope of this model, the nodes get identified with the sites of the
    lattice and the edges signal the neighbor status of two sites. If the
    corresponding nodes are connected, the sites will be neighbors.\\
    All here mentioned graph types are \emph{proximity graphs}.
    The edges of these graphs connect nodes which are by some metric near
    to each other.
    Hence they are suited to generalize problems defined on regular
    lattices with nearest neighbor relationships.
    In this thesis the distance is always determined by the Euclidean
    metric in two dimensions, though in principle every metric in any
    dimension can be used.\\

    The Gabriel graph (GG) \cite{Gabriel1969} is a subgraph of the
    Delaunay triangulation. Two nodes \(i\) and \(j\) with distance

@@

    The Relative Neighborhood graph (RNG) \cite{Toussaint1980} is a

@@

                exsist, of RNG (hatched region) and GG (cross hatched region)

@@

    there are \(N(N-1)\) pairs and for each \(N-2\) nodes to test. So

@@

    To reduce the complexity one can first create a Delaunay
    Triangulation in complexity \(O (N \log N)\)
    \cite{RNGCell} and test the criterion for each edge, because
    the Delaunay triangulation is a supergraph of both. But the
    implementation of a Delaunay triangulation algorithm is not trivial
    and the generation of the graphs is not time critical in the scope

@@

    the criterion for nodes which are near to the lune and abort if

@@

    near the lune one can subdivide the area in \emph{cells} and save
    for each cell a list with nodes lying inside like presented in
    \cite{RNGCell} and sketched in Fig.\ \ref{fig:cell}.

@@

    Now it is just necessary to test the nodes in the cells which

>>>>

    The most common definition of the Ising model, to which will be referred to
    as the \emph{square lattice Ising model}, is a square lattice with
    edge length \(L\) and \(N=L^2\) sites with periodic boundary conditions.
    Note that \(N\) always refers to the number of sites in this thesis.
    Each site has a magnetic moment, which is called the spin. Each

@@

        H = - \sum_{\avg{i,j}}J_{ij}s_{i}s_{j}.

@@

    the coupling constant. If \(J_{ij} > 0 \ \forall \avg{i,j}\) the model resembles a ferromagnet.
    For the square lattice Ising model \(J_{ij} = 1 \ \forall \avg{i,j}\) applies.\\

@@

    thesis is that the sites of the square lattice are displaced introducing
    geometric disorder resulting in a non regular graph structure.

@@

    In the following \(\sigma\) will also be called \emph{disorder parameter}.
    If we take "nearest neighbor" in the Euclidean meaning, most sites
    will only have one nearest neighbor after the

@@

    would destroy the planar character of the model.

@@

    so that for a given configuration of points the resulting graph is an
    instance of a proximity graph. The coupling constant \(J\) gets
    identified with edge weights. \(J\) will be changed to depend on the
    geometric distance between the connected pair of sites. More precise,
    the weight of an edge is \(E_{ij} = J_{ij} = \exp (\alpha (1-d_{ij}))\)
    where \(d_{ij}\) is the Euclidean
    distance between the nodes \(i\) and \(j\). Following Ref.\ \cite{Lima2000}
    the free parameter \(\alpha\) is set to \(\alpha = 0.5\).

@@

            On this torus, for a more clear presentation, the underlying
            graph exhibits a height to width ratio of 1:4. At 1:1 the torus would cut

@@

            Also the shades are of course only a guide to the eye.

@@

    already studied on a Delaunay triangulation. Ref.\ \cite{Janke1994} examines
    this for constant \(J\) and Ref.\ \cite{Lima2000} for \(J_{ij} = e^{-\alpha d_{ij}}\).

@@

    exponents, as can be expected since the underlying graph structure can
    be embedded in 2D. Hence, due to universality it should exhibit the same
    critical exponents as the basic square lattice Ising ferromagnet.

\subsection{Gabriel- and Relative Neighborhood Graph}
\label{ssec:graphtypes}
    A graph \(G(V,E)\) is a set of nodes \(V\) and edges \(E\). In the
    scope of this model, the nodes get identified with the sites of the
    lattice and the edges signal the neighbor status of two sites. If the
    corresponding nodes are connected, the sites will be neighbors.\\
    All here mentioned graph types are \emph{proximity graphs}.
    The edges of these graphs connect nodes which are by some metric near
    to each other.
    Hence they are suited to generalize problems defined on regular
    lattices with nearest neighbor relationships.
    In this thesis the distance is always determined by the Euclidean
    metric in two dimensions, though in principle every metric in any
    dimension can be used.\\
    \paragraph{The Delaunay triangulation} (DT) TODO!!
    \paragraph{The Gabriel graph} (GG) \cite{Gabriel1969} is a subgraph of the
    DT. Two nodes \(i\) and \(j\) with distance

@@

    \paragraph{The Relative Neighborhood graph} (RNG) \cite{Toussaint1980} is a

@@

                exist, of RNG (hatched region) and GG (cross hatched region)

@@

    there are \(N(N-1)\) pairs and for each \((N-2)\) nodes to test. So

@@

    To reduce the complexity one can first create a DT in complexity \(O (N \log N)\)
    \cite{RNGCell} and test the connectedness for each edge contained
    therein, because the DT is a supergraph of both.
    For the construction of the DT for a given set of points one might
    use existing software libraries, e.g.\ the qHull\footnote{http://www.qhull.org} library.
    However, generation of the graphs is not time critical in the scope

@@

    the connectedness for nodes which are near to the lune and abort if

@@

    near the lune one can subdivide the area in \(M\) \emph{cells} of size
    \(L_c \times L_c\) and save for each cell a list with nodes lying
    inside like presented in Ref.\ \cite{RNGCell} and sketched in Fig.\ \ref{fig:cell}.
    The implementation of this thesis uses \(M=N\).

@@

    Given the discretized cell structure it is just necessary to test the nodes in the cells which
